### 2017-04-06 paper/evaluation.tex
merge

%\begin{figure}[t!]
\begin{wrapfigure}{r}[0pt]{0pt}%0.525\linewidth
    \footnotesize
    \centering
    \setlength{\tabcolsep}{2.5pt}
%    \resizebox{.48\textwidth}{!}{%
    \begin{tabular}{lrr}
        \toprule
        \emph{Subject} & \emph{\# of modules} & \emph{Mode}\\%
        \midrule%
        \Comment{BounceStorage }Chaos\Comment{ HTTP Proxy} & 1 &  \ParClassSeqMeth{}\\%
        \Comment{Apache }Flink & 74 & \ForkSeq{} \\%        
        \Comment{JenkinsCI }Gerrit\Comment{ Trigger Plugin} & 1 & \ForkSeq{}\\%
        \Comment{Spotify }Helios & 8 & \ForkSeq{}\\%
        Javaslang & 3 & \ParClassParMeth{}\\%
        Jcabi\Comment{ Github} & 1 & \ParClassParMeth{}\\%        
        \Comment{Hazelcast }Jet & 7 & \ForkSeq{}\\%
        \Comment{Apache Logging }Log4J2 & 25 & \ForkSeq{}\\%
        \Comment{Jankotek }MapDB & 1 & \ParClassParMeth{}\\%        
        \bottomrule%
    \end{tabular}
%    }
    \caption{Subjects with parallel test execution enabled by
    default.}
    \label{tab:freqmodes-dynamic}
    %\end{figure}
\end{wrapfigure}      
Figure~\ref{tab:freqmodes-dynamic} shows the projects we idendified
where parallelism is enabled by default.  Column ``\emph{Subject}''
indicates the name of the project, column ``\emph{\# of modules}''
indicates the number of detected modules containing tests, and column
``\emph{Mode}'' shows the parallel configuration mode used in those
modules (see Section~\ref{sec:modes}).  It came as a surprise the
observation that no project used distinct configurations in their
modules.\Mar{$\leftarrow$ table indicates this.  please confirm.}
Considering our set of \numMedLong{} projects, we found that only
\textbf{\numProjectsPar{}} of those projects had parallelism enabled
by default, with only configurations \ParClassSeqMeth{},
\ParClassParMeth{}, and \ForkSeq{} being used.  Configurations
\ParClassParMeth{} and \ForkSeq{} were the most popular among these
cases.  Note that these results under-approximate real usage of
parallelism as we used default parameters in our scripts to spawn the
Maven process.  That decision could prevent execution of particular
test modules.  To complement this experiemnt, we also looked in
\emph{pom.xml} files for the presence of parallelism.
%\begin{figure}[h!]
%    \centering
%    \includegraphics[width=0.32\textwidth]{plots/barplot-modes-dynamic.pdf}
%    \caption{\label{fig:freqmodes-dynamic}\Fix{fix
%    caption}Distribution of parallel modes identified dynamically in a
%    subset of \numProjectsPar{} projects.  A project may have support
%    to more than one parallel mode. Also, a project may run only a
%    subset tests in parallel by default.}
%\end{figure}

\subsubsection{Presence of parallelism configurations in build files}
\label{sec:rqC-2}

\Mar{----------------------- estou aqui}

To extend our previous investigation, we evaluated the presence of
parallelism configuration in several build files (\ie, \pomf{} files)
from the same sample set of \numMedLong{} projects. The rationale is
that some projects may have support to parallel execution but it is
deactivated by default (\ie, require additional arguments before the
test execution). For instance, on Maven, it is possible to create
multiple configurations in the same build file and activate any of
them dynamically.  To reduce the space of \Fix{x} build files, we
considered a subset of \Fix{y} build files with the following
criteria: the given \pomf{} file has \CodeIn{maven-surefire-plugin}
plugin declaration and it has an explicit \CodeIn{configuration}
declaration. We automatically parsed these \Fix{y} files and we used
the same approach to classify the parallelization mode from the parsed
configurations as in our previous analysis. In an initial analysis, we
noticed that some projects do not declare a static value in some
parameters, instead, these projects inform the actual value in the
command line. To handle this case, we computed all valid values for
the parameter.

@@

Figure~\ref{tab:inspection-table} summarizes our results.
\Fix{The first column indicates the group of projects according to
their time cost.  The second column indicates the number of build
files per group.  The last column indicates the ratio of projects with
parallelization settings.  From the \numMedLong{} subjects, we found
\pomMedLong{} \pomf{} files.  The \numPomMatched{} configurations are
distributed across \numProjectsPar{} projects from our sample.}

% \emph{From these results we found that $\sim$51\% of medium and
% long-running projects do not use parallel features to run test
% suites.}\Mar{please make it consistent with research
% question}\Mar{explain this is over(under)-estimated...}
\begin{figure}[ht!]
    \centering
    \resizebox{.48\textwidth}{!}{%
    \begin{tabular}{llcl}
        \toprule
        Group & Subject & \# of modules & Mode\\%
        \midrule%
        Long   &JenkinsCI Gerrit Trigger Plugin& 1 & \ForkSeq\\%
        Medium &Bouncestorage Chaos Http Proxy & 1 & C2\\%
        Medium &Javaslang & 1 & C3\\%
        Medium &Apache Flink & 1 &\ForkSeq\\%
        Medium &Apache Logging Log4J2 & 3 & \ForkSeq{}\\%
        Medium &Google Cloud Platform DataflowJavaSDK & 1 & C3\\%
        Medium &Hazelcast Jet & 1 & \ForkSeq\\%
        Medium &Jankotek MapDB & 1 & C3\\%
        Medium &Mapstruct & 1 & \ForkSeq\\%
        Medium &Spotify Helios & 3 & \ForkSeq\\%
        \bottomrule%
    \end{tabular}}
    \caption{Subjects with parallelization configurations in build files.}
    \label{tab:inspection-table}
\end{figure}
%% \begin{figure}[ht!]
%%     \centering
%%     \begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}ccc}
%%         \toprule
%%         \multirow{2}{*}{Group} %1st row, 1st cell
%%             & \multirow{2}{*}{\# \pomf{}}
%% 	    & \# \pomf{} matched\\
%%         % 2nd row - empty cell
%%             & % empty cell
%%             & / total\\%
%%         \midrule%
%% 	Long   & \numPomLong{} & 4 / \numLong{}\\%
%% 	Medium & \numPomMed{} & 6 / \numMed{}\\%
%%         \midrule%
%%         Total % last row, first cell
%%             & \pomMedLong{}
%%             & \numProjectsPar{} / \numMedLong{}\\%
%%         \bottomrule%
%%     \end{tabular*}
%%     \caption{Presence of parallelization settings in build files: the
%%     first column indicates the group of projects according to their
%%     time cost; the second column is the subset of files with parallelization
%%     keywords; the last column indicates the ratio of projects with
%%     parallelism support.}
%%     \label{tab:inspection-table} 
%% \end{figure}
%% \Jbc{rework this... $\rightarrow$} From the \numProjectsPar{} projects
%% identified above, we investigated further the \numPomMatched{}
%% build files with parallel settings.  We analyzed the support and
%% distribution of parallel modes from this subset of projects. To
%% calculate the distribution of parallel modes, we considered only the
%% presence of the mode in at least one of the project settings.  Recall
%% that a build file may contain more than one parallel setting and a
%% project may contain several sub-modules with build files.  In case the
%% value of a parallel option is resolved dynamically (\eg, via
%% command-line argument or system variable) we compute all modes related
%% to the option. For instance, depending on the value, the
%% \CodeIn{parallel} option can be \Seq{} (\CodeIn{none}),
%% \ParClassSeqMeth{} (\CodeIn{classes}), \SeqClassParMeth{},
%% (\CodeIn{methods}), and \ParClassParMeth{} (\CodeIn{all}).
%% Figure~\Fix{fig:freqmodes-static} summarizes our findings.
%% \Fix{Missing conclusion: Fork the most used configuration}
%% \begin{figure}[h!]
%%     \centering
%%     \includegraphics[width=0.32\textwidth]{plots/barplot-modes-static.pdf}
%% 	\caption{\label{fig:freqmodes-static}\Luis{This is wrong, it
%% 	should be \textbf{CF0} instead of \textbf{CL0}}Distribution of parallel modes
%%     identified statically in a subset of \numProjectsPar{} projects.
%%     A project may have support to more than one parallel mode.}
%% \end{figure}

\begin{center}
\fbox{
  \begin{minipage}{8cm}
      \textit{Answering \numRQC{}:}~\emph{Overall, results indicate
      that parallelism features are underused by default given that
      only \percentParallel{} of projects  (5 out of \numMedLong)
      supports parallel execution of test suites by default and
      \Fix{12\%} (\Fix{10} out of \numMedLong{}) require additional
      parameters to enable these features.}
  \end{minipage}
}
\end{center}

\begin{itemize}
	\item \textbf{\RQE{}}
\end{itemize}

>>>>

%\begin{figure}[t!]
\begin{wrapfigure}{r}[0pt]{0pt}%0.525\linewidth
    \footnotesize
    \centering
    \setlength{\tabcolsep}{2.5pt}
%    \resizebox{.48\textwidth}{!}{%
    \begin{tabular}{lrr}
        \toprule
        \emph{Subject} & \emph{\# of modules} & \emph{Mode}\\%
        \midrule%
        \Comment{BounceStorage }Chaos\Comment{ HTTP Proxy} & 1 &  \ParClassSeqMeth{}\\%
        \Comment{Apache }Flink & 74 & \ForkSeq{} \\%        
        \Comment{JenkinsCI }Gerrit\Comment{ Trigger Plugin} & 1 & \ForkSeq{}\\%
        \Comment{Spotify }Helios & 8 & \ForkSeq{}\\%
        Javaslang & 3 & \ParClassParMeth{}\\%
        Jcabi\Comment{ Github} & 1 & \ParClassParMeth{}\\%        
        \Comment{Hazelcast }Jet & 7 & \ForkSeq{}\\%
        \Comment{Apache Logging }Log4J2 & 25 & \ForkSeq{}\\%
        \Comment{Jankotek }MapDB & 1 & \ParClassParMeth{}\\%        
        \bottomrule%
    \end{tabular}
%    }
    \caption{Subjects with parallel test execution enabled by
    default.}
    \label{tab:freqmodes-dynamic}
    %\end{figure}
\end{wrapfigure}      
Figure~\ref{tab:freqmodes-dynamic} shows the projects we idendified
where parallelism is enabled by default.  Column ``\emph{Subject}''
indicates the name of the project, column ``\emph{\# of modules}''
indicates the number of detected modules containing tests, and column
``\emph{Mode}'' shows the parallel configuration mode used in those
modules (see Section~\ref{sec:modes}).  It came as a surprise the
observation that no project used distinct configurations in their
modules. Considering our set of \numMedLong{} projects, we found that
only \textbf{\numProjectsPar{}} of those projects had parallelism
enabled by default, with only configurations \ParClassSeqMeth{},
\ParClassParMeth{}, and \ForkSeq{} being used.  Configurations
\ParClassParMeth{} and \ForkSeq{} were the most popular among these
cases.  Note that these results under-approximate real usage of
parallelism as we used default parameters in our scripts to spawn the
Maven process.  That decision could prevent execution of particular
test modules.  
%\begin{figure}[h!]
%    \centering
%    \includegraphics[width=0.32\textwidth]{plots/barplot-modes-dynamic.pdf}
%    \caption{\label{fig:freqmodes-dynamic}\Fix{fix
%    caption}Distribution of parallel modes identified dynamically in a
%    subset of \numProjectsPar{} projects.  A project may have support
%    to more than one parallel mode. Also, a project may run only a
%    subset tests in parallel by default.}
%\end{figure}

\subsubsection{Presence of parallelism configurations in build files}
\label{sec:rqC-2}
Given the inherent limitation of dynamic monitoring to find evidence
of parallelism, we also looked for indication of parallelism in
\emph{pom.xml} files\Comment{ in the same sample set of \numMedLong{}
  projects}. In principle, some projects can use parallel execution
although that is only activated when developers pass certain
parameters to the build process. For instance, it is possible to
create in Maven multiple configurations in the same build file and
select dynamically which one should be used.

\Mar{I have no clue what you mean here... Why should I care about
  space on build files?  I thought we did this check
  for every project and module.  $\rightarrow$}\Fix{To reduce the
  space of \Fix{x} build files, we considered a subset of \Fix{y}
  build files with the following criteria}: the given \pomf{} file has
\CodeIn{maven-surefire-plugin} plugin declaration and it has an
explicit \CodeIn{configuration} declaration. We automatically parsed
these \Fix{y} files and we used the same approach to classify the
parallelization mode from the parsed configurations as in our previous
analysis. In an initial analysis, we noticed that some projects do not
declare a static value in some parameters, instead, these projects
inform the actual value in the command line. To handle this case, we
computed all valid values for the parameter.

We found ten projects that use parallelism using this methodology.
Compared to the set of projects listed in Figure~\ref{tab:speedup}, we
found two new projects, namely: \CodeIn{Google Cloud\Comment{
    Platform} DataflowJavaSDK} (using configuration C3) and
\CodeIn{Mapstruct} (using configuration \ForkSeq{}).  Curiously, we
also found that project \CodeIn{Jcabi} was not detected using this
methodology.  That happened because \Mar{please, explain why}.
Considering the static and dynamic methods together, we found a total
of 11 distinct projects using parallelism, corresponding to the union
of the two subject sets.

\begin{center}
\fbox{
  \begin{minipage}{8cm}
      \textit{Answering \numRQC{}:}~\emph{Results indicate that test
        suite parallelization is underused.  Overall, only
        \percentParallel{} of costly projects (11 out of \numMedLong)
        use parallelism.}
  \end{minipage}
}
\end{center}

@@

%% Figure~\ref{tab:inspection-table} summarizes our results.
%% \Fix{The first column indicates the group of projects according to
%% their time cost.  The second column indicates the number of build
%% files per group.  The last column indicates the ratio of projects with
%% parallelization settings.  From the \numMedLong{} subjects, we found
%% \pomMedLong{} \pomf{} files.  The \numPomMatched{} configurations are
%% distributed across \numProjectsPar{} projects from our sample.}

%% % \emph{From these results we found that $\sim$51\% of medium and
%% % long-running projects do not use parallel features to run test
%% % suites.}\Mar{please make it consistent with research
%% % question}\Mar{explain this is over(under)-estimated...}
%% \begin{figure}[ht!]
%%     \centering
%%     \resizebox{.48\textwidth}{!}{%
%%     \begin{tabular}{llcl}
%%         \toprule
%%         Group & Subject & \# of modules & Mode\\%
%%         \midrule%
%%         Long   &JenkinsCI Gerrit Trigger Plugin& 1 & \ForkSeq\\%
%%         Medium &Bouncestorage Chaos Http Proxy & 1 & C2\\%
%%         Medium &Javaslang & 1 & C3\\%
%%         Medium &Apache Flink & 1 &\ForkSeq\\%
%%         Medium &Apache Logging Log4J2 & 3 & \ForkSeq{}\\%
%%         Medium &Google Cloud Platform DataflowJavaSDK & 1 & C3\\%
%%         Medium &Hazelcast Jet & 1 & \ForkSeq\\%
%%         Medium &Jankotek MapDB & 1 & C3\\%
%%         Medium &Mapstruct & 1 & \ForkSeq\\%
%%         Medium &Spotify Helios & 3 & \ForkSeq\\%
%%         \bottomrule%
%%     \end{tabular}}
%%     \caption{Subjects with parallelization configurations in build files.}
%%     \label{tab:inspection-table}
%% \end{figure}

%% \begin{figure}[ht!]
%%     \centering
%%     \begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}}ccc}
%%         \toprule
%%         \multirow{2}{*}{Group} %1st row, 1st cell
%%             & \multirow{2}{*}{\# \pomf{}}
%% 	    & \# \pomf{} matched\\
%%         % 2nd row - empty cell
%%             & % empty cell
%%             & / total\\%
%%         \midrule%
%% 	Long   & \numPomLong{} & 4 / \numLong{}\\%
%% 	Medium & \numPomMed{} & 6 / \numMed{}\\%
%%         \midrule%
%%         Total % last row, first cell
%%             & \pomMedLong{}
%%             & \numProjectsPar{} / \numMedLong{}\\%
%%         \bottomrule%
%%     \end{tabular*}
%%     \caption{Presence of parallelization settings in build files: the
%%     first column indicates the group of projects according to their
%%     time cost; the second column is the subset of files with parallelization
%%     keywords; the last column indicates the ratio of projects with
%%     parallelism support.}
%%     \label{tab:inspection-table} 
%% \end{figure}
%% \Jbc{rework this... $\rightarrow$} From the \numProjectsPar{} projects
%% identified above, we investigated further the \numPomMatched{}
%% build files with parallel settings.  We analyzed the support and
%% distribution of parallel modes from this subset of projects. To
%% calculate the distribution of parallel modes, we considered only the
%% presence of the mode in at least one of the project settings.  Recall
%% that a build file may contain more than one parallel setting and a
%% project may contain several sub-modules with build files.  In case the
%% value of a parallel option is resolved dynamically (\eg, via
%% command-line argument or system variable) we compute all modes related
%% to the option. For instance, depending on the value, the
%% \CodeIn{parallel} option can be \Seq{} (\CodeIn{none}),
%% \ParClassSeqMeth{} (\CodeIn{classes}), \SeqClassParMeth{},
%% (\CodeIn{methods}), and \ParClassParMeth{} (\CodeIn{all}).
%% Figure~\Fix{fig:freqmodes-static} summarizes our findings.
%% \Fix{Missing conclusion: Fork the most used configuration}
%% \begin{figure}[h!]
%%     \centering
%%     \includegraphics[width=0.32\textwidth]{plots/barplot-modes-static.pdf}
%% 	\caption{\label{fig:freqmodes-static}\Luis{This is wrong, it
%% 	should be \textbf{CF0} instead of \textbf{CL0}}Distribution of parallel modes
%%     identified statically in a subset of \numProjectsPar{} projects.
%%     A project may have support to more than one parallel mode.}
%% \end{figure}

\begin{itemize}
	\item \numRQE{}. \textbf{\RQE{}}
\end{itemize}
